PANCHANGE_data_preprocessing

### 2020-08-07 PANCHANGE_data_preprocessing2020-07-15.Rmd
reducation of the missingness variable phq

##Missingness of questionnaire items
#PHQ - baseline
PHQ Missingness NAs per person - count
```{r PHQ baseline missingness}
coping.glad.phq.raw.id <- coping.glad.phq.raw.id %>%
  mutate(
    na_per_person_phq_base =         # variable for NA per person
      rowSums(                       
        is.na(.[phq.items_base])
        )
    )

@@

```{r PHQ baseline missingness 8 items}
#We need to identify those who have all items APART from the suicide ideation item - to still include these individuals. 
dat.no.sex.no.gender.no.age.no.ibd <- dat.no.sex.no.gender.no.age.no.ibd %>%
  mutate(
    phq.base_cut_off = 
      if_else(
        na_per_person_phq_base == 1 & 
          is.na(phq.dead_hurting_thoughts_base_numeric), 
        1,
        0
      )
  )

table(dat.no.sex.no.gender.no.age.no.ibd$phq.base_cut_off)
```

#PHQ - retrospective
PHQ Missingness NAs per person - count and percentages
```{r PHQ retrospective missingness}
dat.no.sex.no.gender.no.age.no.ibd <- dat.no.sex.no.gender.no.age.no.ibd %>%
  mutate(
    na_per_person_phq_retro =         # variable for NA per person
      rowSums(                       
        is.na(dat.no.sex.no.gender.no.age.no.ibd # sum the rows that are NA
          [,colnames(dat.no.sex.no.gender.no.age.no.ibd) #for all the rows and columns that are in the PHQ items at retroline
           %in%
             phq.items_retro]
        )
    )
  )

freq(dat.no.sex.no.gender.no.age.no.ibd$na_per_person_phq_retro) #have a look at the frequency

dat.no.sex.no.gender.no.age.no.ibd <- dat.no.sex.no.gender.no.age.no.ibd %>% #NA per person as a percentage
  mutate(
    miss_pc_total_phq_retro = 
      na_per_person_phq_retro/9 # NA per person divided by the amount of items
  )

freq(dat.no.sex.no.gender.no.age.no.ibd$miss_pc_total_phq_retro)
```

```{r PHQ retro missingness 8 items}
#We need to identify those who have all items APART from the suicide ideation item - to still include these individuals. 
dat.no.sex.no.gender.no.age.no.ibd <- dat.no.sex.no.gender.no.age.no.ibd %>%
  mutate(
    phq.retro_cut_off = 
      if_else(
        na_per_person_phq_retro == 1 & 
          is.na(phq.dead_hurting_thoughts_retro_numeric), 
        1,
        0
      )
  )

table(dat.no.sex.no.gender.no.age.no.ibd$phq.retro_cut_off)
```

#PHQ - prepandemic
PHQ Missingness NAs per person - count and percentages
```{r PHQ prepan missingness}
dat.no.sex.no.gender.no.age.no.ibd <- dat.no.sex.no.gender.no.age.no.ibd %>%
  mutate(
    na_per_person_phq_prepan =         # variable for NA per person
      rowSums(                       
        is.na(dat.no.sex.no.gender.no.age.no.ibd # sum the rows that are NA
          [,colnames(dat.no.sex.no.gender.no.age.no.ibd) #for all the rows and columns that are in the PHQ items at prepanline
           %in%
             phq.items_prepan]
        )
    )
  )

freq(dat.no.sex.no.gender.no.age.no.ibd$na_per_person_phq_prepan) #have a look at the frequency

dat.no.sex.no.gender.no.age.no.ibd <- dat.no.sex.no.gender.no.age.no.ibd %>% #NA per person as a percentage
  mutate(
    miss_pc_total_phq_prepan = 
      na_per_person_phq_prepan/9 # NA per person divided by the amount of items
  )

freq(dat.no.sex.no.gender.no.age.no.ibd$miss_pc_total_phq_prepan)
```

```{r PHQ prepan missingness 8 items}
#We need to identify those who have all items APART from the suicide ideation item - to still include these individuals. 
dat.no.sex.no.gender.no.age.no.ibd <- dat.no.sex.no.gender.no.age.no.ibd %>%
  mutate(
    phq.prepan_cut_off = 
      if_else(
        na_per_person_phq_prepan == 1 & 
          is.na(phq.dead_hurting_thoughts_prepan_numeric), 
        1,
        0
      )
  )

table(dat.no.sex.no.gender.no.age.no.ibd$phq.prepan_cut_off)
```

Count for PHQ baseline, retrospective, and prepandemic missingness per person
```{r count of people missing a number of items for PHQ}
#baseline
table(dat.no.sex.no.gender.no.age.no.ibd$na_per_person_phq_base)
#retrospective
table(dat.no.sex.no.gender.no.age.no.ibd$na_per_person_phq_retro)
#pre-pandemic
table(dat.no.sex.no.gender.no.age.no.ibd$na_per_person_phq_prepan)
```

@@

PHQ
The PHQ sum score is set to NA if they have more than one missing item, and it is not the last suicide ideation item (they do not meet the cut off criteria - phq.base_cut_off)
```{r sum score cut offs PHQ base}
dat.no.sex.no.gender.no.age.no.ibd <- dat.no.sex.no.gender.no.age.no.ibd %>%
  mutate(
    phq.sum_score_base_full =
      if_else(na_per_person_phq_base > 1 & phq.base_cut_off == 0,
              NA_real_,
              phq.sum_score_base
              )
  )

>>>>

# PHQ missingness (NAs per person)

## baseline
```{r PHQ baseline missingness}
coping.glad.phq.raw.id <- coping.glad.phq.raw.id %>%
  mutate(
    na_per_person_phq_base =         # variable for NA per person
      rowSums(                       
        is.na(.[phq.items_base])
        )
    )

@@

We need to identify those who have all items APART from the suicide ideation item - to still include these individuals. 
```{r PHQ base missingness 8 items}
coping.glad.phq.raw.id <- coping.glad.phq.raw.id %>%
  mutate(
    phq.missing_only_suicide_item_base = 
      if_else(
        condition = na_per_person_phq_base == 1 & 
          is.na(phq.dead_hurting_thoughts_base_numeric), 
        true = 1,
        false = 0,
        missing = NA_real_
      )
  )

coping.glad.phq.raw.id %>%
  freq(phq.missing_only_suicide_item_base,
       cumul = F)
```

## retrospective
```{r PHQ retrospective missingness}
coping.glad.phq.raw.id <- coping.glad.phq.raw.id %>%
  mutate(
    na_per_person_phq_retro =         # variable for NA per person
      rowSums(
        is.na(.[phq.items_retro])
        )
    )

coping.glad.phq.raw.id %>%
freq(na_per_person_phq_retro, cumul = F) #have a look at the frequency
```

```{r PHQ retro missingness 8 items}
coping.glad.phq.raw.id <- coping.glad.phq.raw.id %>%
  mutate(
    phq.missing_only_suicide_item_retro = 
      if_else(
        condition = na_per_person_phq_retro == 1 & 
          is.na(phq.dead_hurting_thoughts_retro_numeric), 
        true = 1,
        false = 0,
        missing = NA_real_
      )
  )

coping.glad.phq.raw.id %>%
  freq(phq.missing_only_suicide_item_retro,
       cumul = F)
```

## prepandemic
```{r PHQ prepan missingness}
glad.phq.raw.id <- glad.phq.raw.id %>%
  mutate(
    na_per_person_phq_prepan =         # variable for NA per person
      rowSums(                       
        is.na(.[phq.items_prepan])
        )
    )

glad.phq.raw.id %>%
  freq(na_per_person_phq_prepan,
       cumul = F)
```

```{r PHQ prepan missingness 8 items}
glad.phq.raw.id <- glad.phq.raw.id %>%
  mutate(
    phq.missing_only_suicide_item_prepan = 
      if_else(
        condition = na_per_person_phq_prepan == 1 & 
          is.na(phq.dead_hurting_thoughts_prepan_numeric), 
        true = 1,
        false = 0,
        missing = NA_real_
      )
  )

glad.phq.raw.id %>%
  freq(phq.missing_only_suicide_item_prepan,
       cumul = F)
```

@@

PHQ
The PHQ sum score is set to NA if they have more than one missing item, and it is not the last suicide ideation item (they do not meet the cut off criteria - phq.missing_only_suicide_item)
```{r sum score cut offs PHQ base}
dat.no.sex.no.gender.no.age.no.ibd <- dat.no.sex.no.gender.no.age.no.ibd %>%
  mutate(
    phq.sum_score_base_full =
      if_else(na_per_person_phq_base > 1 & phq.missing_only_suicide_item == 0,
              NA_real_,
              phq.sum_score_base
              )
  )
